\papertitle

### 2025-03-21 src/tex/ms.tex
[showyourwork] overleaf sync

To outline the circular averaging approach from Bai and Breen, known in directional statistics as the extrinsic mean,~\cite{hotz_intrinsic_2011} we consider a simple one-dimensional periodic system, where the cell ranges from \num{0} to \num{1}. 
This system is analogous to a fractional coordinate system. 
A group of $N$ particles can be described with two vectors, one for the positions, $\bm{x}$, and another for the masses, $\bm{m}$. 
Each particle $i$ is then projected onto the two dimensions of a unit circle, $\bm{\xi}$ and $\bm{\zeta}$, where, 
%
\begin{equation}
    \xi_i = \cos(2\pi x_i),
\end{equation}
%
and 
%
\begin{equation}
    \zeta_i = \sin(2\pi x_i).
\end{equation}
The average of the vectors $\bm{\xi}$ and $\bm{\zeta}$ are then found, weighted by the masses, $m$, of $i$,
%
\begin{equation}
    \bar{\xi} = \frac{\sum_{i=1}^{N}m_i\xi_i}{\sum_{i=1}^{N}m_i},
\end{equation}
% 
and
%
\begin{equation}
    \bar{\zeta} = \frac{\sum_{i=1}^{N}m_i\zeta_i}{\sum_{i=1}^{N}m_i}.
\end{equation}
% 
$\bar{\xi}$ and $\bar{\zeta}$ represent the coordinates of a point on the plane formed by the circle. 
The projection process is then reversed for the weighted average circular coordinates.
Practically, this is achieved using the 2-argument arctangent, 
%
\begin{equation}
    \bar{x} = \frac{\text{atan}2(-\bar{\zeta}, -\bar{\xi}) + \pi}{2\pi},
\end{equation}
%
to give the centre of mass position.
In the case of an orthogonal multi-dimensional system, this method can be repeated independently for each dimension.

>>>>

To outline the circular averaging approach from Bai and Breen, known in directional statistics as the extrinsic mean,~\cite{hotz_intrinsic_2014} we consider a simple one-dimensional periodic system, where the cell ranges from \num{0} to \num{1}. 
This system is analogous to a fractional coordinate system. 
A group of $N$ particles can be described with two vectors, one for the positions, $\bm{x}$, and another for the masses, $\bm{m}$. 
Each particle $i$ is then projected onto the two dimensions of a unit circle, $\bm{\xi}$ and $\bm{\zeta}$, where, 
%
\begin{equation}
    \xi_i = \cos(2\pi x_i),
\end{equation}
%
and 
%
\begin{equation}
    \zeta_i = \sin(2\pi x_i).
\end{equation}
The average of the vectors $\bm{\xi}$ and $\bm{\zeta}$ are then found, weighted by the masses, $m$, of $i$,
%
\begin{equation}
    \bar{\xi} = \frac{\sum_{i=1}^{N}m_i\xi_i}{\sum_{i=1}^{N}m_i},
\end{equation}
% 
and
%
\begin{equation}
    \bar{\zeta} = \frac{\sum_{i=1}^{N}m_i\zeta_i}{\sum_{i=1}^{N}m_i}.
\end{equation}
% 
$\bar{\xi}$ and $\bar{\zeta}$ represent the coordinates of a point on the plane formed by the circle. 
The projection process is then reversed for the weighted average circular coordinates.
Practically, this is achieved using the 2-argument arctangent, 
%
\begin{equation}
    \bar{x} = \frac{\text{atan}2(-\bar{\zeta}, -\bar{\xi}) + \pi}{2\pi},
\end{equation}
%
to give the centre of mass position.
In the case of an orthogonal multi-dimensional system, this method can be repeated independently for each dimension.
